---
title: "Handling Time in Python"
author: "Fan Yin"
date: '2021-01-19T21:13:14-05:00'
output: html_document
categories: Time Series
tags: ["Python Time Data"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<script src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.5/MathJax.js?config=TeX-AMS_CHTML.js"></script>

In this blog, I'll summarize a generic way to find out the number of seconds, minutes, days between two time points in Python using the functions provided in the powerful pandas and numpy libraries. Thanks to the brilliant R package reticulate, I am able to run Python codes in R and write this blog comfortably using R markdown.

## Preparing the environment

### Loading required R libraries
```{r library, include=FALSE}
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE)
  try(sapply(pkg, require, character.only = TRUE), silent = TRUE)
}
packages <- c("dplyr","ggplot2","reticulate")
ipak(packages)
```

### Set up the R-Python environment
```{r python, message=FALSE}
library(reticulate)
# use_python(Sys.which("python"))
# use_condaenv("py3.7", required = TRUE)
use_condaenv("r-reticulate")
```

### Install required Python libraries
```{r python library installation, message=FALSE}
py_install("pandas") # install pandas if it is not available
py_install("numpy")
```

### Import required Python libraries
```{python load library, message = FALSE}
import pandas as pd
import numpy as np
```

## Timestamps in pandas
In pandas, we can access the current time easily via
```{python pandas time 1, message = FALSE}
timestamp_now_UTC = pd.Timestamp.now(tz="UTC")
```

We can also construct an arbitrary time by setting up the year, month, day, hour, minute and even second
```{python pandas time 2, message = FALSE}
timestamp_UTC = pd.Timestamp(year=2020, month = 8, day = 1, hour = 1, minute = 1, second = 25, tz="UTC")
```

We can find out the difference between the above two timestamps by 
```{python pandas time diff, message = FALSE}
# find out the difference
timestamp_diff = timestamp_now_UTC - timestamp_UTC
print(timestamp_diff)
```

However, pandas does not provide us an easy way to convert this difference into a number representing this difference in the unit of days, hours and minutes. To this end, we might want to leverage some functions in numpy.


## Convert time difference into numbers using numpy

If we want to know the difference in the unit of hours
```{python numpy time diff h, message = FALSE}
#
UNIT = "h"
timestamp_diff_h = timestamp_diff.to_numpy().astype("timedelta64" + "[" + UNIT + "]").astype(np.int32)
print(timestamp_diff_h)
#
# timestamp_diff_h_float = timestamp_diff.to_numpy().astype("timedelta64" + "[" + UNIT + "]").astype(np.float32)
# print(timestamp_diff_h_float)
```

If we want to know the difference in the unit of seconds
```{python numpy time diff s, message = FALSE}
#
UNIT = "s"
timestamp_diff_s = timestamp_diff.to_numpy().astype("timedelta64" + "[" + UNIT + "]").astype(np.int32)
print(timestamp_diff_s)
#
# timestamp_diff_s_float = timestamp_diff.to_numpy().astype("timedelta64" + "[" + UNIT + "]").astype(np.float32)
# print(timestamp_diff_s_float)
```


